# Human geography lab: Annex 4

## US states and regions

I will explore different packages that could help with mapping geographical data, specifically for the USA. At first we need to load necessary packages to execute the maps for this lecture.

```{r}
pacman::p_load(usmap, tidyverse)
```

-   Alternatively you can use the pacman package manager for loading the necessary packages for this lecture.

```{txt}
#| code-fold: false
#| warning: false
#| message: false
pacman::p_load(usmap, maps, mapdata, sf, plotly, tidyverse, mapview, leaflet, tmap, tmaptools, tigris)
```

## Using `usmap` package

-   This is probably the most continent package to plot a USA. map along with some data. Plot all states of the USA to create an empty map.

```{r}
#| code-fold: false
#| warning: false
#| message: false
plot_usmap(regions = "states") + 
  labs(title = "United States",
       subtitle = "This is a map of the United States.") + 
  theme(panel.background=element_blank())
```

-   Besides states, we could also plot all counties of the USA.

```{r}
#| code-fold: false
#| warning: false
#| message: false
plot_usmap(regions = "counties") + 
  labs(title = "U.S.A. counties",
       subtitle = "This is a map of the United States.") + 
  theme(panel.background=element_blank())
```

-   `usmap` also provides many different regions to select

```{r}
#| code-fold: false
#| warning: false
#| message: false
plot_usmap(include = .south_region, exclude = c("VA"), labels = TRUE)
```

-   `usmap` also provides north regions map

```{r}
#| code-fold: false
#| warning: false
#| message: false
plot_usmap(include = .south_region, exclude = c("VA"), labels = TRUE)
```

-   We can also select certain states by using *state abbreviations*

```{r}
#| code-fold: false
#| warning: false
#| message: false
plot_usmap(include = c("CT", "ME", "MA", "NH", "VT")) +
  labs(title = "New England Region") +
  theme(panel.background = element_rect(color = "black"))
```
