# Human geography lab: Annex 2

I will explore different packages that could help with mapping geographical data, specifically for the USA. At first we need to load necessary packages to execute the maps for this lecture.

```{r}
pacman::p_load(maps, usmap, tidyverse, DT)
```

### EXPANSION OF WALMART

Walmart has become one of the largest retail multinational companies in the world. Table shows the names and descriptions of variables in the Walmart store opening data, `walmart.csv`.

```{r}
walmart <- read.csv("data/csv/walmart.csv")
## red = Wal-MartStore, green = SuperCenter, blue = DistributionCenter
walmart$storecolors <- NA # create an empty vector
walmart$storecolors[walmart$type == "Wal-MartStore"] <-
    rgb(red = 1, green = 0, blue = 0, alpha = 1/3)
```

The DT package (an interface to the DataTables JavaScript library) can display R matrices or data frames as interactive HTML tables that support filtering, pagination, and sorting.

```{r}
library(DT)
datatable(walmart)
```

The data set contains three different types of stores, represented by the variable type. Wal-MartStore represents a standard Walmart store, whereas SuperCenter is a standard Walmart store as well as a full supermarket. Walmart Supercenters often include pharma- cies, garden shops, car service centers, and other specialty centers. We also plot DistributionCenter data, representing stores that distribute food and goods to standard Walmart stores and Supercenters.

```{r}
walmart$storecolors[walmart$type == "SuperCenter"] <-
    rgb(red = 0, green = 1, blue = 0, alpha = 1/3)
walmart$storecolors[walmart$type == "DistributionCenter"] <-
    rgb(red = 0, green = 0, blue = 1, alpha = 1/3)
## larger circles for DistributionCenter
walmart$storesize <- ifelse(walmart$type == "DistributionCenter", 1, 0.5)
```

Finally, we create a map and add Walmart store locations to it. We also include a legend using the legend() function. To use this function, we specify the location of the legend by setting the x and y coordinates and provide a vector of legend texts as the legend argument.

```{r}
## map with legend
maps::map(database = "state")
points(walmart$long, walmart$lat, col = walmart$storecolors,
       pch = 19, cex = walmart$storesize)
legend(x = -120, y = 32, bty = "n",
       legend = c("Walmart", "Supercenter", "Distribution center"),
       col = c("red", "green", "blue"), pch = 19, # solid circles
       pt.cex = c(0.5, 0.5, 1)) # size of circles
```

```{r}
walmart.map <- function(data, date) {
    walmart <- subset(data, subset = (opendate <= date))
    maps::map(database = "state")
    points(walmart$long, walmart$lat, col = walmart$storecolors,
pch = 19, cex = walmart$storesize)
}
```

```{r}
walmart$opendate <- as.Date(walmart$opendate)
walmart.map(walmart, as.Date("1974-12-31"))
title("1975")
```

```{r}
walmart.map(walmart, as.Date("1984-12-31"))
title("1985")
```

```{r}
walmart.map(walmart, as.Date("1994-12-31"))
title("1995")
```

```{r}
walmart.map(walmart, as.Date("2004-12-31"))
title("2005")
```
