# Human geography lab_2

The following packages will be required for this chapter:

```{r}
#| warning: false
#| message: false
#| code-fold: false
pacman::p_load(tidyverse, maps)
```

### World map

-   The `maps` package contains data for the world, USA, France, Italy and New Zealand. It contains information such as latitude, longitude, group, region, etc. like this.

-   First call the world map data

-   You can see the data structure by using `head` command.

```{r message=FALSE, warning=FALSE}
#| code-fold: false
world_map <- map_data("world")
head(world_map)
```

-   By using the `ggplot2` boundaries can be drawn with `geom_polygon()` with latitude and longitude as shown in @fig-mymap-12.1

```{r fig-mymap-12.1, message=FALSE, warning=FALSE}
#| code-fold: false
#| fig.cap: "World countries"
#| fig.height: 4
world_map %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "black", size = 0.1) 
```

### Japan map

-   If we limit the data with `filter()`, we can also draw a map of specific countries.

-   For instance, we will filter the data for Japan and draw a map as shown in @fig-mymap-12.2

```{r fig-mymap-12.2, message=FALSE, warning=FALSE}
#| code-fold: false
#| fig.cap: "Japan"
#| fig.height: 4
world_map %>% 
  filter(region == "Japan") %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "black", size = 0.1)
```

-   We will change the color of Japan map and draw a map as shown in @fig-mymap-12.3

```{r fig-mymap-12.3, message=FALSE, warning=FALSE}
#| code-fold: false
#| fig.cap: "Japan"
#| fig.height: 4
world_map %>% 
  filter(region == "Japan") %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "blue", colour = "black", size = 0.1)
```

-   We will filter the data for China and draw a map as shown in @fig-mymap-12.4

```{r fig-mymap-12.4, message=FALSE, warning=FALSE}
#| code-fold: false
#| fig.cap: "China"
#| fig.height: 4
world_map %>% 
  filter(region == "China") %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "red", colour = "black", size = 0.1)
```

