# Human geography lab_2

The following packages will be required for this chapter:

```{r}
#| warning: false
#| message: false
#| code-fold: false
pacman::p_load(tidyverse, maps, countrycode, rnaturalearth, sf)
```

### World map

-   The `maps` package contains data for the world, USA, France, Italy and New Zealand. It contains information such as latitude, longitude, group, region, etc. like this.

-   First call the world map data

-   You can see the data structure by using `head` command.

```{r message=FALSE, warning=FALSE}
#| code-fold: false
world_map <- map_data("world")
head(world_map)
```

-   By using the `ggplot2` boundaries can be drawn with `geom_polygon()` with latitude and longitude as shown in @fig-mymap-12.1

```{r fig-mymap-12.1, message=FALSE, warning=FALSE}
#| code-fold: false
#| fig.cap: "World countries"
#| fig.height: 4
world_map %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "black", size = 0.1) 
```

### Japan map

-   If we limit the data with `filter()`, we can also draw a map of specific countries.

-   For instance, we will filter the data for Japan and draw a map as shown in @fig-mymap-12.2

```{r fig-mymap-12.2, message=FALSE, warning=FALSE}
#| code-fold: false
#| fig.cap: "Japan"
#| fig.height: 4
world_map %>% 
  filter(region == "Japan") %>% 
  ggplot(aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "lightgray", colour = "black", size = 0.1)
```

-   We use free data on world map: Natural Earth

-   Using `rnaturalearth` package, we get the data necessary to draw a world map

-   Name the data world_map

```{r message=FALSE, warning=FALSE}
#| code-fold: false
global_map <- ne_countries(scale = "small", returnclass = "sf")
```

-   Show the list of variables included global_map

```{r message=FALSE, warning=FALSE}
#| code-fold: false
names(global_map)
```

-   We see 169 variables
-   Check the class of global_map
-   If you want to see the data of GDP for each country, use following code:

```{r}
#| code-fold: false
global_map %>% 
select(sovereignt, gdp_md, gdp_year) %>% 
DT::datatable()
```

-   World is a GeoDataFrame with the following columns:

`pop_est`: Contains a population estimate for the country

`continent`: The country's continent

`name_en`: The country's name in English

`iso_a3`: The country's 3 letter abbreviation (we made this the index)

`gdp_md`: A country's GDP in million dollar

`geometry`: A POLYGON for each country (we will learn more about these soon)

```{r}
#| code-fold: false
class(global_map)
```

-   global_map has two classes: sf & data.frame

## Global GDP map

```{r message=FALSE, warning=FALSE}
#| code-fold: false
global_map %>% 
  ggplot() +
  geom_sf(aes(fill = gdp_md)) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(fill = "GDP") +
  theme_minimal(base_family = "HiraKakuProN-W3") +
  theme(legend.position = "bottom")
```

-   Let's change the color and the position of legend.

```{r message=FALSE, warning=FALSE}
#| code-fold: false
global_map %>% 
  ggplot() +
  geom_sf(aes(fill = gdp_md)) +
  scale_fill_gradient(low = "gray", high = "blue") +
  labs(fill = "GDP_2019") +
  theme_minimal(base_family = "HiraKakuProN-W3") +
  theme(legend.position = "left")
```

## Regional GDP map

-   GDP of European countries in 2019

```{r message=FALSE, warning=FALSE}
#| code-fold: false
global_map %>% 
  ggplot() +
  geom_sf(aes(fill = gdp_md)) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(fill = "GDP_Europe") +
  coord_sf(xlim = c(-10, 45), ylim = c(35, 60)) +
  theme_minimal(base_family = "HiraKakuProN-W3") +
  theme(legend.position = "left")
```

-   GDP of African countries in 2019

```{r message=FALSE, warning=FALSE}
#| code-fold: false
global_map %>% 
  ggplot() +
  geom_sf(aes(fill = gdp_md)) +
  scale_fill_gradient(low = "white", high = "green") +
  labs(fill = "GDP_Africa") +
  coord_sf(xlim = c(-30, 60), ylim = c(-40, 40)) +
  theme_minimal(base_family = "HiraKakuProN-W3") +
  theme(legend.position = "left")
```

## World and continents with `sf`

We need the following packages to deal with this lecture:

```{r}
#| code-fold: false
library(sf)
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
```

-   Alternatively, you can use the `pacman` package manager to load necessary R packages from below:

```{txt}
pacman::p_load(sf, tidyverse, rnaturalearth)
```

The above command will install all dependencies required to reproduce the entire lecture

-   Use the package `sf`

-   We will use the world dataset provided by `spData`, to show what sf objects are and how they work. `world` is an 'sf data frame' containing spatial and attribute columns, the names of which are returned by the function names() (the last column in this example contains the geographic information):

```{r}
#| code-fold: false
#| warning: false
#| message: false
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
names(world)
```

-   Make sophistocated and professional looking map of World

```{r}
#| code-fold: false
#| warning: false
#| message: false
world.plot = ggplot() +               
  geom_sf(data = world)
world.plot 
```

-   We will begin to get to know the basics of geographic data in R by using the "world" dataset in the package.

-   First let's see what information is in the "world" dataset. Get the names of the variables in the world dataset.

-   Filter the population and GDP data from World map for better understanding the data structure.

-   Plot the GDP percapita map of the world:

-   Prepare Africa map by highlighting Zambia

```{r}
#| code-fold: false
#| warning: false
#| message: false
# Select just the continent of Africa
africa <- world %>% 
  filter(region_un %in% "Africa")
```

-   Check the type of data

```{r}
#| code-fold: false
#| warning: false
#| message: false
class(africa)
names(africa)
```

-   Make professional looking map of `Asia`

```{r}
#| code-fold: false
#| warning: false
#| message: false
asia <- world[which(world$continent == "Asia"),]
asia.plot = ggplot() +               
  geom_sf(data = asia)  
asia.plot 
```

-   Make professional looking map of `Africa`

```{r}
#| code-fold: false
#| warning: false
#| message: false
africa <- world[which(world$continent == "Africa"),]
africa.plot = ggplot() +               
  geom_sf(data = africa)  
africa.plot 
```

-   Make professional looking map of `Europe`

```{r}
#| code-fold: false
#| warning: false
#| message: false
europe <- world[which(world$continent == "Europe"),]
ggplot(europe) +
  geom_sf() +
  coord_sf(xlim = c(-25,50), ylim = c(35,70), expand = FALSE)
```

-   Make professional looking map of `Oceania`

```{r}
#| code-fold: false
#| warning: false
#| message: false
oceania <- world[which(world$continent == "Oceania"),]
ggplot(oceania) +
  geom_sf() +
  coord_sf(xlim = c(105,180), ylim = c(-50,0), expand = FALSE)
```

-   Make professional looking map of `North America`

```{r}
#| code-fold: false
#| warning: false
#| message: false
north_america <- world[which(world$continent == "North America"),]
ggplot(north_america) +
  geom_sf() +
  coord_sf(xlim = c(-180,-15), ylim = c(0,90), expand = FALSE)
```
