# Population density of Bangladesh

-   Load all necessary packages

```{r}
#| message: false
library(tidyverse)
library(sf)
```

-   Alternatively load all necessary packages by using `pacman` package manager

```{txt}
#| code-fold: false
pacman::p_load(tidyverse, stringr, sf, RColorBrewer, leaflet, ggplot2, ggthemes, patchwork, plotly, ggrepel)
```

Load the `shapefile` of Bangladesh admin level 1 which represents the divisions.

```{r, results='hide'}
#| code-fold: false
bd1_shp <- st_read("data/shapefile/bd_adm1_shp/bgd_admbnda_adm1_bbs_20201113.shp")
bd2_shp <- st_read("data/shapefile/bd_adm2_shp/bgd_admbnda_adm2_bbs_20201113.shp")
bd3_shp <- st_read("data/shapefile/bd_adm3_shp/bgd_admbnda_adm3_bbs_20201113.shp")
```

-   Observe the data inside the shape file

```{r, results='hide'}
#| code-fold: false
bd1_shp
```

```{r, results='hide'}
#| code-fold: false
bd2_shp
```

```{r, results='hide'}
#| code-fold: false
bd3_shp
```

-   Check the geographic coordinate system of your shapefile

```{r, results='hide'}
#| code-fold: false
st_crs(bd1_shp)
```

```{r, results='hide'}
#| code-fold: false
st_crs(bd2_shp)
```

```{r, results='hide'}
#| code-fold: false
st_crs(bd3_shp)
```

-   Plot the map and show the name of the divisions

```{r, warning=FALSE}
#| code-fold: false
ggplot(bd1_shp) +
  geom_sf() + 
  geom_sf_text(aes(label = ADM1_EN), size = 2) 
```

```{r, warning=FALSE}
#| code-fold: false
ggplot(bd2_shp) +
  geom_sf() 
```

```{r, warning=FALSE}
#| code-fold: false
ggplot(bd3_shp) +
  geom_sf() 
```

-   Prepare a custom theme to make the visualization of this map better than previous default version

```{r}
#| code-fold: false
theme_custom_map <- function(base_size = 11,
                             base_family = "",
                             base_line_size = base_size / 22,
                             base_rect_size = base_size / 22) {
    theme_bw(base_size = base_size, 
             base_family = base_family,
             base_line_size = base_line_size) %+replace%
        theme(
            axis.title = element_blank(), 
            axis.text = element_blank(),
            axis.ticks = element_blank(),
            complete = TRUE
        )
}
```

```{r, warning=FALSE}
#| code-fold: false
ggplot(bd1_shp) +
  geom_sf() +
  geom_sf_text(aes(label = ADM1_EN), size = 2) +
  theme_custom_map()
```

```{r, warning=FALSE}
#| code-fold: false
ggplot(bd2_shp) +
  geom_sf() +
   theme_custom_map()
```

```{r, warning=FALSE}
#| code-fold: false
ggplot(bd3_shp) +
  geom_sf() +
   theme_custom_map()
```

-   Import population data of all divisions of Bangladesh

```{r}
#| code-fold: false
bd1_pop <- read.csv("data/csv/bd_pop_adm1_2022.csv")
bd2_pop <- read.csv("data/csv/bd_pop_adm2_2022.csv")
bd3_pop <- read.csv("data/csv/bd_pop_adm3_2022.csv")
```

-   Merge the shapefile and csv data-set

```{r, results='hide'}
#| code-fold: false
bd1_shp_pop <- merge(bd1_shp, bd1_pop, by=c("ADM1_PCODE"))
bd1_shp_pop <- st_make_valid(bd1_shp_pop)
glimpse(bd1_shp_pop)
```

```{r, results='hide'}
#| code-fold: false
bd2_shp_pop <- merge(bd2_shp, bd2_pop, by=c("ADM2_PCODE"))
bd2_shp_pop <- st_make_valid(bd2_shp_pop)
glimpse(bd2_shp_pop)
```

```{r, results='hide'}
#| code-fold: false
bd3_shp_pop <- merge(bd3_shp, bd3_pop, by=c("ADM3_PCODE"))
bd3_shp_pop <- st_make_valid(bd3_shp_pop)
glimpse(bd3_shp_pop)
```

-   Draw the map ADM1

```{r}
#| code-fold: false
#| warning: false
ggplot(bd1_shp_pop) +
    geom_sf(aes(fill=F_TL), color = "white", lwd = 0.07) +
    geom_sf_text(aes(label = ADM1_EN), size = 2) +
    scale_fill_viridis_c(option = "magma", name = "Population")
```

-   By loading additional package, we can draw for informative maps

```{txt}
#| code-fold: false
ggplot(bd1_shp_pop) +
    geom_sf(aes(fill=F_TL), color = "white", lwd = 0.07) +
    geom_sf_text(aes(label = ADM1_EN), size = 2) +
    scale_fill_viridis_c(option = "magma", name = "Population") +
    theme_map() + 
    theme(legend.direction="horizontal") +
     labs(title = "Divisions of Bangladesh", 
         subtitle = "Population density of Bangladesh", 
         caption = "Source of data: BBS")
```

-   Draw the map ADM2

```{r}
#| code-fold: false
#| warning: false
ggplot(bd2_shp_pop) +
    geom_sf(aes(fill=F_TL), color = "white", lwd = 0.07) +
    #geom_sf_text(aes(label = ADM2_EN), size = 2) +
    scale_fill_viridis_c(option = "magma", name = "Population")
```

-   By loading additional package, we can draw for informative maps

```{txt}
#| code-fold: false
ggplot(bd2_shp_pop) +
    geom_sf(aes(fill=F_TL), color = "white", lwd = 0.07) +
    #geom_sf_text(aes(label = ADM2_EN), size = 2) +
    scale_fill_viridis_c(option = "magma", name = "Population") +
    theme_map() + 
    theme(legend.direction="horizontal") +
     labs(title = "Districts of Bangladesh", 
         subtitle = "Population density of Bangladesh", 
         caption = "Source of data: BBS")
```

-   Draw the map ADM3

```{r, warning=FALSE}
#| code-fold: false
#| warning: false
ggplot(bd3_shp_pop) +
    geom_sf(aes(fill=F_TL), color = "white", lwd = 0.07) +
    #geom_sf_text(aes(label = ADM2_EN), size = 2) +
    scale_fill_viridis_c(option = "magma", name = "Population")
```

-   By loading additional package, we can draw for informative maps

```{txt}
#| code-fold: false
ggplot(bd3_shp_pop) +
    geom_sf(aes(fill=F_TL), color = "white", lwd = 0.07) +
    #geom_sf_text(aes(label = ADM2_EN), size = 2) +
    scale_fill_viridis_c(option = "magma", name = "Population") +
    theme_map() + 
    theme(legend.direction="horizontal") +
     labs(title = "Upazila/Thana of Bangladesh", 
         subtitle = "Population density of Bangladesh", 
         caption = "Source of data: BBS")
```
